---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = TRUE
)

library(opendap.catalog)
library(terra)
```


<!-- badges: start -->
[![Dependencies](https://img.shields.io/badge/dependencies-7/33-orange?style=flat)](#)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://choosealicense.com/licenses/mit/)
[![Website deployment](https://github.com/mikejohnson51/opendap.catalog/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/mikejohnson51/opendap.catalog/actions/workflows/pkgdown.yaml)
[![LifeCycle](man/figures/lifecycle/lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R CMD Check](https://github.com/mikejohnson51/opendap.catalog/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mikejohnson51/opendap.catalog/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

One of the biggest challenges with Earth System and spatial research is extracting data. These challenges include not only finding the source data but then downloading, managing, and extracting the partitions critical for a given task. 

Services exist to make data more readily available over the web but introduce new challenges of identifying subsets, working across a wide array of standards (e.g. non-standards), all without alleviating the challenge of finding resources. 

In light of this, `opendap.catolog` provides three primary services. 

#### 1. Generalized access of local and remote NetCDF data with `dap()` for space (XY) and Time (T) subsets

```{r}
dap <- dap(URL = "https://cida.usgs.gov/thredds/dodsC/bcsd_obs", 
           AOI = AOI::aoi_get(state = "FL"), 
           startDate = "1995-01-01")

str(dap, max.level = 1)
```

```{r, echo = FALSE}
plot(rast(dap))
```

```{r}
file <- '/Users/mjohnson/Downloads/NEXGDM_srad_2020_v100.nc'
utils:::format.object_size(file.size(file), "auto")

dap = dap(URL = file, 
               AOI = AOI::aoi_get(state = "FL"), 
               startDate = "2020-01-01", endDate = "2020-01-05")
```

```{r, echo = FALSE}
plot(rast(dap))
```

#### 2. A catalog of `r nrow(opendap.catalog::params)` web resources that can be accessed programatically (as of 02/2022)

```{r}
dplyr::glimpse(opendap.catalog::params)
```

For use in other applications (e.g. stars proxy, geoknife, or python) this catalog is available as a JSON artifact [here](https://mikejohnson51.github.io/opendap.catalog/cat_params.json).

```{r, eval  = FALSE}
read_json('https://mikejohnson51.github.io/opendap.catalog/cat_params.json', 
          simplifyVector = TRUE)
```


### (3) The ability to pass catalog elements to the generalized toolsets to aid data discovery and complicated edge cases:

```{r}
# Find MODIS PET in Florida for January 2010
(
  dap = dap(
    catolog = dplyr::filter(params, 
                            id == 'MOD16A2.006', 
                            varname == 'PET_500m'),
    AOI = AOI::aoi_get(state = "FL"),
    startDate = "2010-01-01",
    endDate   = "2010-01-31"
  )
)
```

```{r, echo = FALSE}
plot(dap)
```




