---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = TRUE
)

library(opendap.catalog)
library(terra)
```


<!-- badges: start -->
[![Dependencies](https://img.shields.io/badge/dependencies-5/26-orange?style=flat)](#)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://choosealicense.com/licenses/mit/)
[![Website deployment](https://github.com/mikejohnson51/opendap.catalog/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/mikejohnson51/opendap.catalog/actions/workflows/pkgdown.yaml)
[![LifeCycle](man/figures/lifecycle/lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R CMD Check](https://github.com/mikejohnson51/opendap.catalog/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mikejohnson51/opendap.catalog/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

<div id="top"></div>

<br />
<div align="center">

<h1 align="center"><strong>OpenDap</strong> Catalog</h1>

  <p align="center">
    <a href="https://mikejohnson51.github.io/opendap.catalog/"><strong>« Explore the Docs »</strong></a>
    <br />
    <br />
    <a href="https://mikejohnson51.github.io/opendap.catalog/cat_params.json">Data Catalog</a>
    ·
    <a href="">R Interface</a>
    ·
    <a href="https://github.com/mikejohnson51/opendap.catalog/issues">Request Feature</a>
  </p>
</div>

<hr>

One of the biggest challenges with Earth System and spatial research is extracting data. These challenges include not only finding the source data but then downloading, managing, and extracting the partitions critical for a given task. 

Services exist to make data more readily available over the web but introduce new challenges of identifying subsets, working across a wide array of standards (e.g. non-standards), all without alleviating the challenge of finding resources. 

In light of this, `opendap.catolog` provides three primary services. 

<hr>


#### 1. Generalized space (XY) and Time (T) subsets for _remote_ and _local_ NetCDF data with `dap()` 

> remote 

```{r}
dap <- dap(URL = "https://cida.usgs.gov/thredds/dodsC/bcsd_obs", 
           AOI = AOI::aoi_get(state = "FL"), 
           startDate = "1995-01-01")

str(dap, max.level = 1)
```

```{r, echo = FALSE}
plot(rast(dap))
```

> local 

```{r}
file <- '/Users/mjohnson/Downloads/NEXGDM_srad_2020_v100.nc'
utils:::format.object_size(file.size(file), "auto")

dap = dap(URL = file, 
          AOI = AOI::aoi_get(state = "FL"), 
          startDate = "2020-01-01", endDate = "2020-01-05")
```

```{r, echo = FALSE}
plot(rast(dap))
```

#### 2. A catalog of `r formatC(nrow(opendap.catalog::params),big.mark = ",",digits = 0,format = "f")` web resources (as of `r format(Sys.Date(), format = "%m/%Y")`)

```{r}
dplyr::glimpse(opendap.catalog::params)
```

For use in other applications (e.g. [stars proxy](https://github.com/r-spatial/stars/pull/499), [geoknife](https://github.com/USGS-R/geoknife), [climateR](https://github.com/mikejohnson51/climateR) or python/go/Rust applciations) this catalog is available as a JSON artifact [here](https://mikejohnson51.github.io/opendap.catalog/cat_params.json).

```{r, eval  = FALSE}
read_json('https://mikejohnson51.github.io/opendap.catalog/cat_params.json', 
          simplifyVector = TRUE)
```

With  `r formatC(nrow(opendap.catalog::params),big.mark = ",",digits = 0,format = "f")` web resources documented, there are simply too many resources to search through by hand unless you know exactly what you want. This  voids the possibility of serendipitous discovery. So, we have added a generally fuzzy search tool to help discover datasets.

Say you want to find what resoruces there are for monhtly snow water equivilent? `search` and `search_summary` can help: 

```{r}
search("monthly swe") 
```
We could also search for daily precipitation in MACA for scenario RCP45 and summarize the results: 
 
```{r}
search("daily precipitation maca rcp45") |> 
  search_summary()
```
Overall we see there are 20 model/ensemeble members that produced MACA daily rainfall.

### (3) The ability to pass catalog elements to the generalized toolsets:

```{r}
# Find MODIS PET in Florida for January 2010
dap = dap(
    catolog = search("MOD16A2.006 PET"),
    AOI = AOI::aoi_get(state = "FL"),
    startDate = "2010-01-01",
    endDate   = "2010-01-31"
  )
```

```{r, echo = FALSE}
plot(dap)
```




