---
title: "OpenDAP catalog"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OpenDAP catalog}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning  = FALSE,
  message  = FALSE,
  comment = "#>"
)
```

```{r setup}
library(opendap.catalog)
library(dplyr)
library(terra)
```

# Intro

The initial Vignette highlighted how to extract subsets of data when the source data set was known. A second goal of `opendap.catalog` is to provide an evolving catalog of available data resources with the information needed to access them. 
As of `r format(Sys.Date(), format = "%m/%Y")` there are `r nrow(opendap.catalog::params)` datasets identified and stored in `opendap.catalog::params`. In these there are `r length(unique(opendap.catalog::params$id))` unique data sources, and `r length(unique(opendap.catalog::params$varname))` unique variables.

```{r}
dplyr::glimpse(opendap.catalog::params)
```
For use in other applications (e.g. [stars proxy](https://github.com/r-spatial/stars/pull/499), [geoknife](https://github.com/USGS-R/geoknife), [climateR](https://github.com/mikejohnson51/climateR) or python/go/Rust applciations) this catalog is available as a JSON artifact [here](https://mikejohnson51.github.io/opendap.catalog/cat_params.json).

```{r, eval  = FALSE}
read_json('https://mikejohnson51.github.io/opendap.catalog/cat_params.json', simplifyVector = TRUE)
```

# Contents

| Name      | Description | Details |
|---------- | ----------- | -------- | 
| id        | `opendap.catalog`  unique identifier                   | | 
| grid.id   | `opendap.catalog`  unique  grid identifier             | See opendap.catalog::grids |
| URL       | Location of resource                                   | |
| tiled     | Is the resource temporally (T) or spatially (XY) tiled | See Vignette 3|
| variable  | name of the variable in data description               | |
| varname   | name of variable in data set description               | |
| long_name | NetCDF Long Name attribute                             | |
| units     | NetCDF Long Name units attribute                       | |
| model     | Model Type/version                                     | Often GCM |
| ensemble  | Model ensemble member                                  | |
| scenario  | Model scenario                                         | Often RCPs |
| T_name    | Name of time dimension                                 | |
| duration  | start and end date of data (start/end)                 | |
| interval  | time step between duration                             | |
| nT        | number of time slices                                  | |

# Example

```{r}
search <- filter(params, 
                 grepl("NLDAS", id),
                 grepl("FOR", model),
                 grepl("potential evaporation", long_name),
                 scenario == 'M.2.0',
                 grepl("month", interval))


glimpse(search)
```

Catalog item rather then URL!

```{r}
tx_2021_pet = dap(catolog = search, 
                  AOI = AOI::aoi_get(state = "TX"),
                  startDate = "2021-01-01", 
                  endDate = "2021-12-31")
```

```{r, echo = FALSE}
plot(tx_2021_pet$potevap_M.2.0)
```

# Grid Relation

```{r}
glimpse(grids)
```

| Name | Description | Details |
|----- | ----------- | -------- | 
| grid.id       | `opendap.catalog`  unique  grid identifier                      | |
| X_name        | X dimension name in data set                                    | |
| Y_name        | Y dimension name in data set                                    | |
| X1            | The first X value in the X_name array                           | |
| Xn            | The last X value in the X_name array                            | |
| Y1            | The first Y value in the X_name array                           | |
| Yn            | The last Y value in the X_name array                            | |
| resX          | Resolution of X_name array                                      | |
| resY          | Resolution of Y_name array                                      | |
| ncols         | Number of columns in the XY array                               | |
| nrows         | Number of rows in the XY array                                  | |
| proj          | Data set projection                                             | |
| toptobottom   | It is organized in left to right,top to bottom pixel order.     | |
| tile          | If the dataset has a tiled XY partition, what is the tile name? | |

```{r}
glimpse(merge(search, grids))
```

Like `parms`, `grids` attributes can also be accessed as a JSON:

```{r, eval  = FALSE}
read_json('https://mikejohnson51.github.io/opendap.catalog/cat_grids.json', simplifyVector = TRUE)
```


# Challenge

```{r}
(search <- filter(params, 
                  grepl("MOD16A2.006", id), 
                  grepl("PET_500m", varname)))

glimpse(merge(search, grids))
```


