---
title: "LCMAP Examples"
author: "Mike Johnson"
date: "3/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      error = FALSE, 
                      message = FALSE)

source("../private/aws.R")
library(terra)
```

# General Notes

We are looking for efficient, remote-access to LCMAP (and NLCD) data that is not reliant on a service like WCS or REST APIs.

The goal is to be able to access each file at a single non-authenticated end point, and that each file is easily discoverable (either online or via a catalog). The semi-new VSI capabilities in GDAL offer a way to achieve this using the existing LCMAP resources, however they are not stored effiecently for this purpose given they are zipped.

The aim is to show what is possible with what is currently available and - hopefully - provide rationale for a minimally-invasive step towards more efficient data access.

# Use Case

Say we want 2020 land cover data for a county in Colorado (Larimer). We will use this as a respective AOI.

```{r}
AOI <- AOI::aoi_get(state = "CO", county = "Larimer")
```

This county is ~6,825 sqkm with a bounding box area of ~8,700 sqkm. This is approximetly 10,600,000 LCMAP 30 meter cells.

```{r, echo = FALSE}
co <- AOI::aoi_get(state = c("CO"))
{
  plot(co$geometry)
  plot(AOI$geometry, add = TRUE, col = "red")
}
```

# Examples

### VSI over zipped resources

We _can_ backdoor into the existing resources using a fairly complex request that requires knowing the location of the zipped directory and the included file names.

These _could_ be cataloged...

```{r}
system.time({
  r <- rast("/vsizip/{/vsicurl/https://edcintl.cr.usgs.gov/downloads/sciweb1/shared/lcmap/public/full_extent_downloads/version_12/primary-landcover_conus_year_data/LCMAP_CU_2020_V12_LCPRI.zip}/LCMAP_CU_2020_V12_LCPRI.tif")
  r2 <- crop(r, project(vect(AOI), crs(r)))
})
```

```{r, echo = FALSE}
plot(r2)
```

While this _does_ meet the conditions outlined above it is fairly slow...

### s3 VSI

As an alternative, I downloaded, unzipped, and moved the data to a s3 bucket (not public sorry). Here we can use a much simplier vsi call.

```{r}
system.time({
  r <- rast("/vsis3/formulations-dev/spatial-grids/LCMAP_CU_2020_V12_LCPRI.tif")
  r2 <- crop(r, project(vect(AOI), crs(r)))
})
```


Overall this approach takes ~2% of the time needed to extracted data from the zipped version.

### Is a VRT better?

We have been working to show the advantages of VRT formats with the [National Map Data](https://mikejohnson51.github.io/opendap.catalog/articles/usgs_vrts.html). Since LCMAP is not tiled I do not expect improvements working over the VRT, but, we can check:

```{r}
system.time({
  r <- rast("/vsis3/formulations-dev/spatial-grids/LCMAP_CU_2020_V12_LCPRI.vrt")
  r2 <- crop(r, project(vect(AOI), crs(r)))
})
```

### COG

It is not clear if the tif files provided by LCMAP are COGs (the xml file suggests they are?). Either way, we also convert the LCMAP TIF to a uncompressed COG for completeness, 
This creates a _large_ file - blew up from 1.2 GB to 22GB w/o compression. This might prove prohibitive for the LCMAP team given there are 35 years * 10 varibable (~7,700 GB!!)

I added this file to the same s3 bucket and we can time access to it:

```{r, eval = FALSE, echo = FALSE}
system("gdal_translate /Users/mjohnson/Downloads/LCMAP_CU_2020_V12_LCPRI/LCMAP_CU_2020_V12_LCPRI.tif /Users/mjohnson/Downloads/LCMAP_CU_2020_V12_LCPRI/LCMAP_CU_2020_V12_LCPRI_cog.tif -of COG")
```

```{r}
system.time({
  r <- rast("/vsis3/formulations-dev/LCMAP_CU_2020_V12_LCPRI_cog.tif")
  r2 <- crop(r, project(vect(AOI), crs(r)))
})
```

### CLI, Python and more...

So far we have shown access uisng r-spatial tools, however any language that has a GDAL binding can be used...

```{r}
sf::write_sf(AOI, "aoi.gpkg")

system.time({
  system("gdalwarp -cutline aoi.gpkg -crop_to_cutline /vsis3/formulations-dev/spatial-grids/LCMAP_CU_2020_V12_LCPRI.vrt aoi.tif")
})
```


```{r, echo = FALSE}
plot(rast("aoi.tif"))

# Unlink
unlink("aoi.tif")
unlink("aoi.gpkg")
```


# End objective

So why is this needed? Here we provide two basic use cases that support USGS, NOAA and community goals.

### Zonal Stats

One of objectives we are trying to achieve with improved LCMAP access is the ability to rapidly generate summary attribute descriptors over large sets of spatial units.

The "pseudo" AOI in this example is all NHD catchments in Larimer county - we can get this with `nhdplusTools`:

```{r}
nhd <- nhdplusTools::get_nhdplus(AOI = AOI, realization = "catchment")
```

In total this provides us with `r nrow(nhd)` catchments, covering the `r ncell(r2)` LCMAP cells. To compute these, we can utilize the `zonal` toolset:

```{r}
system.time({
  sum = zonal::execute_zonal(r2, nhd, FUN = "freq", ID = 'featureid')
})

plot(sum[sum$value == 4, 'percentage'], main = "Percentage LC 4")
plot(sum[sum$value == 2, 'percentage'], main = "Percentage LC 2")
```

**This can also be done straight from web resources!**

Here is a complete remote access summary of LCMAP landcover over NextGEN CAMELS basin. It is in this type of access that the key points outlines in the introduction become critical:

```{r}
system.time({
  sum = zonal::execute_zonal("/vsis3/formulations-dev/spatial-grids/LCMAP_CU_2020_V12_LCPRI.vrt", 
                             sf::read_sf('/vsis3/formulations-dev/CAMELS20/camels_01047000_3321976/spatial/hydrofabric.gpkg', "catchments"), 
                             FUN = "freq", 
                             ID = 'ID')
})

plot(sum[sum$value == 4, 'percentage'], main = "Percentage LC 4")
plot(sum[sum$value == 2, 'percentage'], main = "Percentage LC 2")
```

### Hetergenous Data Access via standard endpoints:

Say we want a collection of data for an area: 

  - clay %
  - land cover 
  - elevation

We can access this data from s3, HTTPS, and from Lynker, Github and Duke. The principles outlined above can help us get there. And while this "solves" out needs I would want the data to remain at the authoritative sources and not redistributed by us:

```{r}
files <- c(
  "/vsis3/formulations-dev/spatial-grids/LCMAP_CU_2020_V12_LCPRI.vrt",
  "/vsicurl/http://mikejohnson51.github.io/opendap.catalog/ned_1_tester.vrt",
  "/vsicurl/http://hydrology.cee.duke.edu/POLARIS/PROPERTIES/v1.0/vrt/clay_mean_0_5.vrt"
)


o <- lapply(1:length(files), function(x) {
  r <- rast(files[x])
  crop(r, project(vect(AOI), crs(r)))
})

```


```{r, echo = FALSE}
plot(o[[1]], main = "LCMAP")
plot(o[[2]], main = "NED 1")
plot(o[[3]], main = "Clay %")
```

# Ask of LCMAP Team?

Is there an already existing access point to the unzipped files? If not might it be considered?

If not, is there any oppisition to us hosting - and - advertising access to unzipped s3 resources in this way via the NOAA NextGen project?

Thanks!
